Stats for the convergence paper

### 2020-01-22 models/paper/convergence_paper_models.Rmd
clay comments

These are the models and stats for the convergence paper, going through it by section number.

@@

Our weapon of choice is typially lme4. But it won't converge with full random effects structure. So we turn to brms.

@@

We definitely want some random slopes (fit1 is worse than the rest). Other than that, the various constellations do not improve the model's expected log pointwise predictive density.

@@

## Bayes factors!

@@

exp((BIC(fit_power2) - BIC(fit_power1))/2)
# 0.01179529
# we favour the null model

@@

The lme4 fit reported in the paper is here.

@@

t1 %>% 
  mutate(
    row_id = 9:1, # watch this
  ) %>% 
  mutate(
    term2 = 
    case_when(
      term == 'verb_baseline_average' ~ 'verb baseline average',
      term == 'bot_lexical_typicalityrandom' ~ 'bot lexical typicality = random',
      term == 'bot_lexical_typicalityatypical' ~ 'bot lexical typicality = atypical',
      term == 'participant_prereg_average' ~ 'participant prereg average',
      term == 'bot_regularisation_rate-40%' ~ 'bot regularisation rate = -40%',
      term == 'bot_regularisation_rate+40%' ~ 'bot regularisation rate = +40%',
      term == 'verb_baseline_average:bot_lexical_typicalityrandom' ~ 'verb baseline average : bot lexical typicality = random',
      term == 'verb_baseline_average:bot_lexical_typicalityatypical' ~ 'verb baseline average : bot lexical typicality = atypical',
      term == '(Intercept)' ~ 'Intercept'
      )
    ) %>% 
  ggplot(aes(x = term2 %>% reorder(row_id), y = estimate)) +
  geom_point() +
  geom_linerange(aes(x = term2, ymin = `2.5%`, ymax = `97.5%`)) +
  geom_hline(aes(yintercept = 0), linetype = 'dashed') +
  xlab('') +
  coord_flip() +
  ggtitle('Estimates and 95% confidence intervals in model for test results')
ggsave('terms_figure.pdf', width = 9, height = 5)

>>>>

These are the models and stats reported in the main text of the convergence paper. We go through the main text by section number and replicate the relevant statistics / figures / models here.

This file references both the main text and the supplementary information. To make things slightly more complicated, the appendix of the main text and the SI overlap. And section lettering is different in the appendix of the main text and the si. Here, we use lettering that matches the si.

@@

Our weapon of choice is typially lme4. But it won't converge with full random effects structure. So we turn to brms. We assume flat priors and a bernoulli error distribution, which is preferred to the binomial distribution in brms for binary outcomes. We compare models using leave-one-out cross-validation.

The Bayesian models are equivalent to the frequentist models reported in the paper, at least conceptually, if not in many other ways.

@@

What do we learn from LOO comparisons?

Fit1 is worse than the rest, so random slopes do improve model fit. Building up the random effect structure for the word random intercept does not improve the model's expected log pointwise predictive density.

@@

## Bayes factors

We could not reject the null on the absence of difference between our typical and atypical (reversed) conditions. Upon reviewer suggestion, we used Bayes factors to test whether this is due to the lack of evidence.

@@

exp((BIC(fit_power2) - BIC(fit_power1))/2)
# 0.01179529
# it is unlikely that there is a difference here

@@

The lme4 fit reported in the paper is here:

@@

t1 %>% 
  mutate(
    row_id = 9:1, # watch this
  ) %>% 
  mutate(
    term2 = 
    case_when(
      term == 'verb_baseline_average' ~ 'verb baseline average',
      term == 'bot_lexical_typicalityrandom' ~ 'bot lexical typicality = random',
      term == 'bot_lexical_typicalityatypical' ~ 'bot lexical typicality = atypical',
      term == 'participant_prereg_average' ~ 'participant prereg average',
      term == 'bot_regularisation_rate-40%' ~ 'bot regularisation rate = -40%',
      term == 'bot_regularisation_rate+40%' ~ 'bot regularisation rate = +40%',
      term == 'verb_baseline_average:bot_lexical_typicalityrandom' ~ 'verb baseline average : bot lexical typicality = random',
      term == 'verb_baseline_average:bot_lexical_typicalityatypical' ~ 'verb baseline average : bot lexical typicality = reversed',
      term == '(Intercept)' ~ 'Intercept'
      )
    ) %>% 
  ggplot(aes(x = term2 %>% reorder(row_id), y = estimate)) +
  geom_point() +
  geom_linerange(aes(x = term2, ymin = `2.5%`, ymax = `97.5%`)) +
  geom_hline(aes(yintercept = 0), linetype = 'dashed') +
  xlab('') +
  coord_flip() +
  ggtitle('Estimates and 95% confidence intervals in model for test results')
ggsave('terms_figure.pdf', width = 9, height = 5)
